---
title: "Data Wrangling Assignment"
author: "Zachary Beneduci"
date: "Mar 18, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load tidyverse and download and unpack the datafile from GitHub:

```{r}
#install.packages("tidyverse")
library(tidyverse)

url <- "https://github.com/noelzach/Reproducibility/raw/main/10_DataWrangling/Microbiome.csv.zip"
download.file(url, "Microbiome.zip")
unzip("Microbiome.zip")
MicrobiomeData <- read.csv("Microbiome.csv", stringsAsFactors = T) # I'll set the string values as factors ahead of time.
```


Q1. Select the following columns.

OTU, SampleID, Abundance, Crop, Compartment, DateSampled, GrowthStage, Treatment, Rep, Fungicide, Kingdom, Phylum, Class, Order, Family, Genus, Species, Taxonomy

```{r}
selection <- c("OTU", "SampleID", "Abundance", "Crop", "Compartment", "DateSampled", "GrowthStage", "Treatment", "Rep", "Fungicide", "Kingdom", "Phylum", "Class", "Order", "Family", "Genus", "Species", "Taxonomy")
MicrobiomeData2 <- select(MicrobiomeData, all_of(selection))
```


Q2. Calculate the mean percent relative abundance across all OTUs by compartment using the pipe operator.

```{r}
head(MicrobiomeData2) # We can see that Abundance is already relative, so just multiply by 100 for percent.
MicrobiomeData2 %>%
  group_by(Compartment) %>%
  mutate(Percent = Abundance*100) %>%
  summarise(Mean = mean(Percent))
```


Q3. Calculate the mean percent relative abundance across all OTUs by compartment and fungicide using the pipe operator.

```{r}
MicrobiomeData2 %>%
  group_by(Compartment, Fungicide) %>%
  mutate(Percent = Abundance*100) %>%
  summarise(Mean = mean(Percent))
```


Q4. Restrict the previous analysis to just the fungal class Dothideomycetes.

```{r}
MicrobiomeData2 %>%
  filter(Class == "Dothideomycetes") %>%
  group_by(Compartment, Fungicide) %>%
  mutate(Percent = Abundance*100) %>%
  summarise(Mean = mean(Percent))
```


Q5. Now do the same analysis but for orders in the class Dothideomycetes.

```{r}
MicrobiomeData2 %>%
  filter(Class == "Dothideomycetes") %>%
  group_by(Fungicide, Order) %>%
  mutate(Percent = Abundance*100) %>%
  summarise(Mean = mean(Percent))
```


Q6. Calculate the standard error for each mean in Q5.

```{r}
MicrobiomeData2 %>%
  filter(Class == "Dothideomycetes") %>%
  group_by(Compartment, Fungicide, Order) %>%
  mutate(Percent = Abundance*100) %>%
  summarise(Mean = mean(Percent),
            n = n(),
            sd.dev = sd(Percent)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>%
  select(Compartment, Fungicide, Order, Mean, std.err)
```


Q7. Select the columns Order, Compartment, Fungicide, and Mean from the output of Q6 and Pivot the dataset to wide format so you have Compartment as column headers.

```{r}
MicrobiomeData2 %>%
  filter(Class == "Dothideomycetes") %>%
  group_by(Compartment, Fungicide, Order) %>%
  mutate(Percent = Abundance*100) %>%
  summarise(Mean = mean(Percent),
            n = n(),
            sd.dev = sd(Percent)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>%
  select(Order, Compartment, Fungicide, Mean, std.err) %>%
  pivot_wider(names_from = Compartment, values_from = Mean)
```


Q8. Recreate this plot using your dataset in Q7.

```{r, fig.height=5, fig.width=5.25}

MicrobiomeData2 %>%
  filter(Class == "Dothideomycetes") %>%
  group_by(Compartment, Fungicide, Order) %>%
  mutate(Percent = Abundance*100) %>%
  summarise(Mean = mean(Percent),
            n = n(),
            sd.dev = sd(Percent)) %>%
  mutate(std.err = sd.dev/sqrt(n)) %>%
  select(Order, Compartment, Fungicide, Mean, std.err) %>%
  pivot_wider(names_from = Compartment, values_from = Mean) %>%
  mutate(Order2 = fct_reorder(Order, desc(-Root))) %>%
  ggplot(aes(x = Order2, y = Root, fill = Fungicide)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  coord_flip()+
  xlab(NULL)+
  ylab("Percent relative abundance") +
  theme_gray(base_size = 12)

```

